---
title: "hw1"
author: "Tariq Anwar"
date: "January 20, 2016"
output: html_document
---

Data Set Source: https://github.com/caesar0301/awesome-public-datasets/blob/master/Datasets/titanic.csv.zip


Nature of Data: The dataset contains data about titanic.

It contains infomation about passengers, Gender, Name, Fare , Survived Passengers etc.


There are total 12 features in the dataset and 891 observations.
              

```{r}
library(ggplot2)
titanic <- read.csv("C:/Users/tariq/Downloads/titanic.csv")
#Summary
summary(titanic)
```


##Question 1.a
###Total Observations ,  Total Features
```{r}
dim(titanic)
#Features of DataSet
names(titanic)
```


##Question 1.b 
###Features Types
```{r}
type<-list()
for(t in names(titanic))
  type[[t]] <- class(titanic[[t]])

 print(type)

```

##Question1.c
###Summaries of each featured is shown above (i.e summary(titanic))

##Question 1.d
###Univariate Plots for all the numeric data:



```{r, echo=FALSE}
#Univariate Plot
quantitative <- sapply(titanic, is.numeric)
for(t in names(titanic[quantitative])){
  hist(titanic[,t],xlab = t, main = paste("Histogram of", t))
}

```


****************************************************************************************************************************************************************************************

#                                    **Question2**
```{r}

set.seed(123)
index <- sample(1:nrow(diamonds), 50) # try a subset first
diamonds2 <- diamonds[index,]
hist(diamonds2$price)
hist(diamonds2$price, breaks=quantile(diamonds2$price))

```

###Question 2.a
We have sample size of 50 diamonds,


The price range is: `r range(diamonds2["price"])`


we can divide the diamonds in different price group of equal size


with  each group having an interval of 2000

Counting the sum of elements in each individual group and storing it in a vector g

```{r}
g<- numeric(length = 0)
g[1]<- (sum(diamonds2["price"]<=2000))
g[2]<- (sum(diamonds2["price"]<=4000 & diamonds2["price"]>2000))
g[3]<- (sum(diamonds2["price"]<=6000 & diamonds2["price"]>4000))
g[4]<- (sum(diamonds2["price"]<=8000 & diamonds2["price"]>6000))
g[5]<- (sum(diamonds2["price"]<=10000 & diamonds2["price"]>8000))
g[6]<- (sum(diamonds2["price"]<=12000 & diamonds2["price"]>10000))
g[7]<- (sum(diamonds2["price"]<=14000 & diamonds2["price"]>12000))
g[8]<- (sum(diamonds2["price"]<=16000 & diamonds2["price"]>14000))

print(g)



```
As we can see , vector g contains same number of elements as 
shown in the given bar graph.


In order to proof the correctness of second histogram , we will calculate each class
width, and that can be caculated by finding the quartile of diamonds "price" parameter
```{r}
quantile(diamonds2$price)
```
Now we can calculate height of each box with the following formula:
$$ h(x) = freq/n*(c(i)-c(i+1)$$
where h = height : 
C(i)= lower boundary of interval :
c(i+1)= upper boundary of interval :
n = total data


```{r}
f<- numeric(length = 0)
f[1]<- (sum(diamonds2["price"]<=1212.75&diamonds2["price"]>=537 ))
f[2]<- (sum(diamonds2["price"]<=2831 & diamonds2["price"]>1212.75))
f[3]<- (sum(diamonds2["price"]<=5724.75 & diamonds2["price"]>2831))
f[4]<- (sum(diamonds2["price"]<=14709 & diamonds2["price"]>5724.75))

#print(f)

#height of each box
h<- numeric(length = 0)
h[1]<-(f[1])/(50*(1212.75-537))
h[2]<-(f[2])/(50*(2831-1212.75))
h[3]<-(f[3])/(50*(5724.75-2831))
h[4]<-(f[4])/(50*(14709-5724.75))



print(h)

```
As we can see the height we calculated manually is almost same as shown in the histogram.


We can cross check the frequency values by below formula
```{r}
w<-numeric(length = 0)
w[1]<- (h[1] * (1212.75-537))*50
w[2]<-(h[2]*(2831-1212.75))*50
w[3]<-(h[3])*(5724.75-2831)*50
w[4]<-(h[4])*(14709-5724.75)*50
print(w)
```
This is the same as the frequency values we calculated above.

If we manually map the height vector and quantile vector it is similar to given histogram.


###Question 2(b):

For the first histogram:


We can count the frequency of first two box , that will give the number of diamonds 
below 4000 

total = freq of 1st box + freq of second box


      = 20+9 => 29
      
  Proportion of diamonds below 4000 = `r 29/50 `    


For second histogram:


Proportion of diamonds with price less than 4000 can be calculated as:

$$ Area of 1st quartile+ Area of second quartile + Area in (4000-2nd quartile boundary)$$

```{r}
val <- (h[1]*(1212.75-537)*50)+(h[2]*(2831-1212.75)*50) + (h[3]*(4000-2831)*50)
```


Proportion of diamonds below 4000 = `r val/50 `